Continous time Markov chain models; \\ extension, simulation and visualisation

### 2014-10-21 chapters/chapter01.tex
improving math

\begin{equation}
P\left\{ N(t+u)-N(u)=k\right\} =P\left\{ N(t)=k\right\}=\frac{e^{-r t}(r t)^{k}}{k!},\; k=0,1,\ldots
\label{eq:poissonDist} 
\end{equation}

@@

\begin{eqnarray}
F_{T_{2}}(t) &=& 1-P\left(T_{2}>t|T_{1}=u\right) \\ \nonumber
& = & 1-P\left(\text{no mutations in }(u,t+u]|T_{1}=u\right) \\ \nonumber 
& = & 1-P\left(N(t)=0\right)=1-e^{-r t}.
\end{eqnarray}

@@

\begin{equation}
\mathbf{V}\times\mathbf{Q}\times\mathbf{V}^{-1}=\left[\begin{array}{cccc}
\lambda_{1} & 0 & \cdots & 0\\
0 & \lambda_{2} & \ddots & \vdots\\
\vdots & \ddots & \ddots & 0\\
0 & \cdots & 0 & \lambda_{K}
\end{array}\right]=\text{diag}(\lambda_{1},\ldots,\lambda_{K})=\mathbf{D} ,
\label{eq:diagonalization}
\end{equation}

@@

\begin{eqnarray}
\mathbf{P}(t) &=& \exp\left(rt\mathbf{Q}\right)  =  \underset{k=0}{\overset{\infty}{\sum}}\left(rt\mathbf{Q}\right)^{k}/k! \\ \nonumber
 &=&  \underset{k=0}{\overset{\infty}{\sum}}\left(\mathbf{V}\times\mathbf{D}\times\mathbf{V}^{-1}\right)^{k}/k! \\ \nonumber
 &=& \mathbf{V}\times\left(\underset{k=0}{\overset{\infty}{\sum}}\mathbf{D}^{k}/k!\right)\times\mathbf{V}^{-1} = \mathbf{V}\times\text{diag}(e^{rt\lambda_{1}},\ldots,e^{rt\lambda_{K}})\times\mathbf{V}^{-1} ,
% \mathbf{P}(t) & = & \exp\left(rt\mathbf{Q}\right)  =  \underset{k=0}{\overset{\infty}{\sum}}\left(rt\mathbf{Q}\right)^{k}/k! 
%  =  \underset{k=0}{\overset{\infty}{\sum}}\left(\mathbf{V}\times\mathbf{D}\times\mathbf{V}^{-1}\right)^{k}/k! \\ \nonumber
% & = & \mathbf{V}\times\left(\underset{k=0}{\overset{\infty}{\sum}}\mathbf{D}^{k}/k!\right)\times\mathbf{V}^{-1} = \mathbf{V}\times\text{diag}(e^{rt\lambda_{1}},\ldots,e^{rt\lambda_{K}})\times\mathbf{V}^{-1} ,
\label{eq:eigen_decomposition}
\end{eqnarray}

@@

\begin{eqnarray}
P\left\{ X(t_{n+1})=i_{n+1}\mid X(t_{n})=i_{n},\ldots, X(t_{0})=i_{0}\right\} & = & \\ \nonumber
P\left\{ X(t_{n+1})=i_{n+1}\mid X(t_{n})=i_{n}\right\} .
\label{eq:markov}
\end{eqnarray}

@@

\begin{eqnarray}
\mathbf{P}\left(t\right) &=& \left[\begin{array}{cccc}
p_{0}(t) & p_{1}(t) & \cdots & p_{1}(t)\\
p_{1}(t) & p_{0}(t) & \ddots & \vdots\\
\vdots & \ddots & p_{0}(t) & p_{1}(t)\\
p_{1}(t) & \cdots & p_{1}(t) & p_{0}(t)
\end{array}\right], \\ \nonumber
&\text{ where }& \ensuremath{\begin{cases}
p_{0}(t)=\frac{1}{4}+\frac{3}{4}e^{-4rt}\\
p_{1}(t)=\frac{1}{4}-\frac{1}{4}e^{-4rt}
\end{cases}}.
\label{eq:jc69Finite}
\end{eqnarray}

@@

%PL: why not use a simpler tree for fewer taxa, not necessarily with dated tips, where you indicate the terms you describe. For example the sample tree for figure 1.10, but without node and branch numbering, and taxon 1, taxon 2,.. as tip labels

@@

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
%
\State $node \gets getRoot\left(\right)$
%
\State $state \gets getState\left(node\right)$
%
\Repeat
%
\If{$\left(hasChildren\left(node\right)\right)$}
%
\ForAll{children}
%
\State $t \gets getDistanceToParent\left(child\right);$
%
\State $ \mathbf{P}\left(t\right) \gets e^{\mathbf{Q}t};$
%
\State $state \gets sample\left(\mathbf{P}\left[ state, \right]\right);$
%
\EndFor
%
\Else \Comment{this is a tip node}
%
\State $node \gets getParent\left(node\right);$
%
\EndIf
%
\Until{$\left(\text{simulated for all nodes}\right)$}
% }
\end{algorithmic}
\caption{
{ \footnotesize 
%PL: Perhaps add: Pseudo code for simulating...
{\bf Simulating an evolutionary process along a phylogeny.} 
When a child node is visited, the state is sampled with conditional probabilities of changing to state $j$ given state $i$ at the parental node.
}% END: footnotesize
}
\label{alg:simulation}
\end{algorithm}

@@

In this thesis we will not discuss the maximum likelihood estimation itself, %PL: good, who wants to consider the data to be fixed anyway..
rather focus on closely related 
%PL: I would add: ".., but philosophically very different,"
methods of Bayesian inference.
The likelihood concept is also required in Bayesian inference and enters the framework as the data component.
We will now discuss how to calculate the likelihood of observing a particular realization of character sequences at the tips of a tree, given a substitution model with fixed parameters.

%---LIKELIHOOD---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{likelihood} 
\caption{
{ \footnotesize 
{\bf  A hypothetical tree for five taxa.} Character data for single site is observed at the tips of the tree.
} % END: footnotesize
}
\label{fig:LIKELIHOOD}
\end{figure}

%PL: I could recommend to align the char's, increase their size, relate them to the tips with horizontal dashed lines, but just increasing their size may do
% also, why not label the external nodes with x1-to x5 as you mention in the text, consistent with the internal models?
% the i indices do not correspond between the nodes and branches for x7 and x8
% make the labels of the axis bigger, or deleted all together as it is not needed.

@@

\begin{eqnarray}
\underset{x_{0}\in \mathcal{E}}{\sum}\;\underset{x_{6}\in \mathcal{E}}{\sum}\;\underset{x_{7}\in \mathcal{E}}{\sum}\;\underset{x_{8}\in \mathcal{E}}{\sum}[\pi_{x_{0}}\cdot p_{x_{0}x_{8}}(t_{7})\cdot p_{x_{0}x_{7}}(t_{8})\cdot p_{x_{8}x_{6}}(t_{6})\cdot p_{x_{8}A}(t_{3}) \cdot \\ \nonumber
p_{x_{7}C}(t_{4})\cdot p_{x_{7}C}(t_{5})\cdot p_{x_{6}T}(t_{1})\cdot p_{x_{6}C}(t_{2})]
\label{eq:likelihoodNaive}
\end{eqnarray}

We can break down the terms in Equation~(\ref{eq:likelihoodNaive}) %for some reason this says 1.22 in my pdf build and not 1.21 as it should be, or is this just due to my TeX environment?
to the probability of observing the data $TCACC$ for the tips and $x_{0},x_{6},x_{7},x_{8}$ for the ancestral nodes (inside the brackets) and the sum over all possible path combinations over the state space $\mathcal{E}$ (outside the brackets).
Term $\pi_{0}$ is simply the probability that the root has character $x_{0}$ and is based on the assumption that the process generating the data has reached its equilibrium.

@@

\begin{eqnarray}
P\left(\theta|\mathbf{X}\right) &=& \frac{P(\theta)\cdot P\left(\mathbf{X}|\theta\right)}{P\left(\mathbf{X}\right)}=\frac{P(\theta)\cdot P\left(\mathbf{X}|\theta\right)}{\int P(\theta)\cdot P\left(\mathbf{X}|\theta\right)d\theta} \\ \nonumber
& \approx & P(\theta)\cdot P\left(\mathbf{X}|\theta\right),
\label{eq:bayesian1}
\end{eqnarray}

@@

\begin{eqnarray}
P\left(\mathbf{\Theta}|\mathbf{X}\right) &=& P\left(\theta,\mathbf{F}|\mathbf{X}\right)=\frac{P\left(\theta,\mathbf{F}\right)\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)}{P\left(\mathbf{X}\right)} \\ \nonumber
&=& \frac{P\left(\theta,\mathbf{F}\right)\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)}{\int\int P(\theta,\mathbf{F})\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)d\theta d\mathbf{F}}
\label{eq:bayesian2}
\end{eqnarray}

@@

% \substack{\text{aaa} \\ \text{bbb}}

\begin{eqnarray}
% \footnotesize{
P\left(\mathbf{F}, \mathbf{Q}, \mathbf{\Phi}|\mathbf{X},\mathbf{Y}\right)\propto 
\underbrace{P(\mathbf{X}|\mathbf{Q}, \mathbf{F})}_{\begin{array}{c}
\substack{\text{Sequence} \\ \text{likelihood}}
\end{array}}\cdot\underbrace{P(\mathbf{Y}|\mathbf{\Phi}, \mathbf{F})}_{\begin{array}{c}
\substack{\text{Trait} \\ \text{likelihood}}
\end{array}}\cdot
\underbrace{P(\mathbf{F})}_{\begin{array}{c}
\substack{\text{Topology} \\ \text{prior}}
\end{array}} \\ \nonumber
\cdot\underbrace{P(\mathbf{Q})}_{\begin{array}{c}
\substack{\text{Sequence substitution} \\ \text{prior}}
\end{array}}\cdot\underbrace{P(\mathbf{\Phi})}_{\begin{array}{c}
\substack{\text{Location exchange} \\ \text{prior}}
\end{array}}
% }
\label{eq:posteriorPhylogeography}
\end{eqnarray}

>>>>

\newpage

\begin{align}
P\left\{ N(t+u)-N(u)=k\right\} & = P\left\{ N(t)=k\right\} \nonumber \\
= \frac{e^{-r t}(r t)^{k}}{k!},\; k=0,1,\ldots
\label{eq:poissonDist} 
\end{align}

@@

\begin{eqnarray}
F_{T_{2}}(t) &=& 1-P\left(T_{2}>t|T_{1}=u\right) \nonumber \\
& = & 1-P\left(\text{no mutations in }(u,t+u]|T_{1}=u\right) \nonumber \\ 
& = & 1-P\left(N(t)=0\right)=1-e^{-r t}.
\end{eqnarray}

@@

\begin{align}
\mathbf{V}\times\mathbf{Q}\times\mathbf{V}^{-1}=\left[\begin{array}{cccc}
\lambda_{1} & 0 & \cdots & 0\\
0 & \lambda_{2} & \ddots & \vdots\\
\vdots & \ddots & \ddots & 0\\
0 & \cdots & 0 & \lambda_{K}
\end{array}\right] %\nonumber \\
= \text{diag}(\lambda_{1},\ldots,\lambda_{K})=\mathbf{D} ,
\label{eq:diagonalization}
\end{align}

@@

\begin{align}
\mathbf{P}(t) & =  \exp\left(rt\mathbf{Q}\right) \nonumber \\
& =   \underset{k=0}{\overset{\infty}{\sum}}\left(rt\mathbf{Q}\right)^{k}/k! \underset{k=0}{\overset{\infty}{\sum}}\left(\mathbf{V}\times\mathbf{D}\times\mathbf{V}^{-1}\right)^{k}/k! \nonumber \\
& =  \mathbf{V}\times\left(\underset{k=0}{\overset{\infty}{\sum}}\mathbf{D}^{k}/k!\right)\times\mathbf{V}^{-1} \nonumber \\
& =  \mathbf{V}\times\text{diag}(e^{rt\lambda_{1}},\ldots,e^{rt\lambda_{K}})\times\mathbf{V}^{-1} ,
\label{eq:eigen_decomposition}
\end{align}

@@

\begin{align}
& P\left\{ X(t_{n+1}) = i_{n+1}\mid X(t_{n})=i_{n},\ldots, X(t_{0})=i_{0}\right\}   \nonumber \\
&= P\left\{ X(t_{n+1}) = i_{n+1}\mid X(t_{n})=i_{n}\right\} .
\label{eq:markov}
\end{align}

@@

\begin{align}
\mathbf{P}\left(t\right) &=& \left[\begin{array}{ccccc}
p_{0}(t) & p_{1}(t) & \cdots & \cdots & p_{1}(t)\\
p_{1}(t) & p_{0}(t) &  &  & \vdots\\
\vdots &  & \ddots & p_{0}(t) & p_{1}(t)\\
p_{1}(t) & \cdots & \cdots & p_{1}(t) & p_{0}(t)
\end{array}\right],  \nonumber \\
&\text{where}& \ensuremath{\begin{cases}
p_{0}(t)=\frac{1}{4}+\frac{3}{4}e^{-4rt}\\
p_{1}(t)=\frac{1}{4}-\frac{1}{4}e^{-4rt}
\end{cases}}.
\label{eq:jc69Finite}
\end{align}

@@

@@

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
%
\State $node \gets getRoot\left(\right)$
%
\State $state \gets getState\left(node\right)$
%
\Repeat
%
\If{$\left(hasChildren\left(node\right)\right)$}
%
\ForAll{children}
%
\State $t \gets getDistanceToParent\left(child\right);$
%
\State $ \mathbf{P}\left(t\right) \gets e^{\mathbf{Q}t};$
%
\State $state \gets sample\left(\mathbf{P}\left[ state, \right]\right);$
%
\EndFor
%
\Else \Comment{this is a tip node}
%
\State $node \gets getParent\left(node\right);$
%
\EndIf
%
\Until{$\left(\text{simulated for all nodes}\right)$}
% }
\end{algorithmic}
\caption{
{ \footnotesize 
{\bf Pseudo code for simulating an evolutionary process along a phylogeny.} 
When a child node is visited, the state is sampled with conditional probabilities of changing to state $j$ given state $i$ at the parental node.
}% END: footnotesize
}
\label{alg:simulation}
\end{algorithm}

@@

In this thesis we will not discuss the maximum likelihood estimation itself, %PL: good, who wants to consider the data to be fixed anyway..
rather focus on closely related 
%PL: I would add: ".., but philosophically very different,"
% FB: difference: ML is an optimization procedure, Bayesian inference is a conditioning. It's a less aggresivee approach 
methods of Bayesian inference.
The likelihood concept is also required in Bayesian inference and enters the framework as the data component.
We will now discuss how to calculate the likelihood of observing a particular realization of character sequences at the tips of a tree, given a substitution model with fixed parameters.

%---LIKELIHOOD---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{likelihood} 
\caption{
{ \footnotesize 
{\bf  A hypothetical tree for five taxa.} Character data for single site is observed at the tips of the tree.
} % END: footnotesize
}
\label{fig:LIKELIHOOD}
\end{figure}

@@

\begin{align}
\underset{x_{0}\in \mathcal{E}}{\sum}\;\underset{x_{6}\in \mathcal{E}}{\sum}\;\underset{x_{7}\in \mathcal{E}}{\sum}\;\underset{x_{8}\in \mathcal{E}}{\sum}[\pi_{x_{0}}\cdot p_{x_{0}x_{8}}(t_{7})\cdot p_{x_{0}x_{7}}(t_{8})\cdot p_{x_{8}x_{6}}(t_{6})\cdot p_{x_{8}A}(t_{3}) \nonumber \\
\cdot p_{x_{7}C}(t_{4})\cdot p_{x_{7}C}(t_{5})\cdot p_{x_{6}T}(t_{1})\cdot p_{x_{6}C}(t_{2})].
\label{eq:likelihoodNaive}
\end{align}

We can break down the terms in Equation~(\ref{eq:likelihoodNaive}) 
to the probability of observing the data $TCACC$ for the tips and $x_{0},x_{6},x_{7},x_{8}$ for the ancestral nodes (inside the brackets) and the sum over all possible path combinations over the state space $\mathcal{E}$ (outside the brackets).
Term $\pi_{0}$ is simply the probability that the root has character $x_{0}$ and is based on the assumption that the process generating the data has reached its equilibrium.

@@

\begin{align}
P\left(\theta|\mathbf{X}\right) &= \frac{P(\theta)\cdot P\left(\mathbf{X}|\theta\right)}{P\left(\mathbf{X}\right)}=\frac{P(\theta)\cdot P\left(\mathbf{X}|\theta\right)}{\int P(\theta)\cdot P\left(\mathbf{X}|\theta\right)d\theta}  \nonumber \\
& \approx  P(\theta)\cdot P\left(\mathbf{X}|\theta\right),
\label{eq:bayesian1}
\end{align}

@@

\begin{eqnarray}
P\left(\mathbf{\Theta}|\mathbf{X}\right) &=& P\left(\theta,\mathbf{F}|\mathbf{X}\right)=\frac{P\left(\theta,\mathbf{F}\right)\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)}{P\left(\mathbf{X}\right)}  \nonumber \\
&=& \frac{P\left(\theta,\mathbf{F}\right)\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)}{\int\int P(\theta,\mathbf{F})\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)d\theta d\mathbf{F}}
\label{eq:bayesian2}
\end{eqnarray}

@@

\begin{align}
P\left(\mathbf{F}, \mathbf{Q}, \mathbf{\Phi}|\mathbf{X},\mathbf{Y}\right) &\propto 
\underbrace{P(\mathbf{X}|\mathbf{Q}, \mathbf{F})}_{\begin{array}{c}
\substack{\text{Sequence} \\ \text{likelihood}}
\end{array}}\cdot\underbrace{P(\mathbf{Y}|\mathbf{\Phi}, \mathbf{F})}_{\begin{array}{c}
\substack{\text{Trait} \\ \text{likelihood}}
\end{array}}\cdot
\underbrace{P(\mathbf{F})}_{\begin{array}{c}
\substack{\text{Topology} \\ \text{prior}}
\end{array}} \nonumber \\
& \cdot \underbrace{P(\mathbf{Q})}_{\begin{array}{c}
\substack{\text{Sequence substitution} \\ \text{prior}}
\end{array}}\cdot\underbrace{P(\mathbf{\Phi})}_{\begin{array}{c}
\substack{\text{Location exchange} \\ \text{prior}}
\end{array}}
\label{eq:posteriorPhylogeography}
\end{align}

### 2014-10-21 chapters/chapter02.tex
last subsection
corrections
improving math

\chapter{Modelling substitution heterogeneity through time\label{chap:epoch}}

@@

Figure~\ref{fig:collapsing} illustrates this action for a branch that spans a single boundary at $T_1$ with $\mathbf{Q}_1$ governing the process before the boundary and $\mathbf{Q}_2$ after the boundary. Letting $X(T_1) = k$ represent the augmented state of the stochastic process at the boundary, we compute
\begin{align}
p_{ij}(t_u, t_v, r_v)
&= \sum_{k} 
\left\{
\exp[ r_v ( T_1 - t_u ) \mathbf{Q}_1]
\right\}_{ik}
\times
\left\{
\exp[ r_v ( t_v - T_1 ) \mathbf{Q}_2]
\right\}_{kj} , 
\end{align}
for all $i,j$, or equivalently in compact matrix form
\begin{align}
\mathbf{P}(t_u, t_v, r_v)
&= 
\exp[ r_v ( T_1 - t_u ) \mathbf{Q}_1]
\times
\exp[ r_v ( t_v - T_1 ) \mathbf{Q}_2] \nonumber \\
&= \mathbf{P}_{1}(t_u, T_1, r_v) \times \mathbf{P}_{2}(T_1, t_v, r_v) ,
\label{matrixProduct}
\end{align}
where 
$\mathbf{P}_{1}(t_u, T_1, r_v)$ and $\mathbf{P}_{2}(T_1, t_v, r_v)$ are shorthand notation used in the figure and again in the next section where it is clear we are considering substitution models for neighboring epochs.

@@

\subsection{Fine-grain parallel implementation}

>>>>

\chapter{Modeling substitution heterogeneity through time\label{chap:epoch}}

@@

Figure~\ref{fig:collapsing} illustrates this action for a branch that spans a single boundary at $T_1$ with $\mathbf{Q}_1$ governing the process before the boundary and $\mathbf{Q}_2$ after the boundary. Letting $X(T_1) = k$ represent the augmented state of the stochastic process at the boundary, we compute

\begin{align}
p_{ij}(t_u, t_v, r_v)
= \sum_{k} 
\left\{
\exp[ r_v ( T_1 - t_u ) \mathbf{Q}_1]
\right\}_{ik} & \nonumber \\
\times 
\left\{
\exp[ r_v ( t_v - T_1 ) \mathbf{Q}_2]
\right\}_{kj} & , 
\end{align}

\noindent
for all $i,j$, or equivalently in compact matrix form

\begin{align}
\mathbf{P}(t_u, t_v, r_v)
&= 
\exp[ r_v ( T_1 - t_u ) \mathbf{Q}_1]
\times
\exp[ r_v ( t_v - T_1 ) \mathbf{Q}_2] \nonumber \\
&= \mathbf{P}_{1}(t_u, T_1, r_v) \times \mathbf{P}_{2}(T_1, t_v, r_v) ,
\label{matrixProduct}
\end{align}
where 
$\mathbf{P}_{1}(t_u, T_1, r_v)$ and $\mathbf{P}_{2}(T_1, t_v, r_v)$ are shorthand notation used in the figure and again in the next section where it is clear we are considering substitution models for neighboring epochs.

@@

\subsection{Fine-grain parallel implementation\label{sub:fine_grain}}

### 2014-10-21 chapters/chapter06.tex
improving math

\footnotesize{
\begin{equation}
q_{ij}^{GY94}=\begin{cases}
\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a synonymous transversion}\\
\kappa\cdot\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a synonymous transition}\\
\omega\cdot\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a non-synonymous transversion}\\
\omega\cdot\kappa\cdot\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a non-synonymous transversion}\\
0 & \text{otherwise}
\end{cases},
\label{eq:gy94}
\end{equation}
} % END: size

@@

{\tiny{
\begin{equation}
q_{ij}^{MG94}=\begin{cases}
\alpha\cdot\kappa\cdot\pi_{n}^{p} & i\rightarrow j\text{ is a synonymous transition from nucleotide \textit{m} to \ensuremath{n} at codon position \ensuremath{p}}\\
\alpha\cdot\pi_{n}^{p} & \text{synonymous transversion from \ensuremath{m} to \ensuremath{n} at position \ensuremath{p}}\\
\beta\cdot\kappa\cdot\pi_{n}^{p} & \text{non-synonymous transition from \ensuremath{m} to \ensuremath{n} at position \ensuremath{p}}\\
\beta\cdot\pi_{n}^{p} & \text{non-synonymous transversion from \ensuremath{m} to \ensuremath{n} at position \ensuremath{p}}\\
0 & \text{otherwise}
% \alpha\cdot\kappa\cdot\pi_{n}^{p} & i\rightarrow j\text{ is a synonymous transition from nucleotide \textit{m} to \ensuremath{n} at codon position \ensuremath{p}.}\\
% \alpha\cdot\pi_{n}^{p} & i\rightarrow j\text{ is a synonymous transversion from nucleotide \ensuremath{m} to \ensuremath{n} at codon position \ensuremath{p}.}\\
% \beta\cdot\kappa\cdot\pi_{n}^{p} & i\rightarrow j\text{ is a non-synonymous transition from nucleotide \ensuremath{m} to \ensuremath{n} at codon position \ensuremath{p}.}\\
% \beta\cdot\pi_{n}^{p} & i\rightarrow j\text{ is a non-synonymous transversion from nucleotide \ensuremath{m} to \ensuremath{n} at codon position \ensuremath{p}.}\\
% 0 & \text{otherwise.}
\end{cases},
\label{eq:mg94}
\end{equation}
} % END: size
}

>>>>

\begin{equation}
\left\{ \mathbf{Q}\right\} _{ij}=\begin{cases}
\pi_{j} & \substack{ \text{ \ensuremath{i\rightarrow j} is a synonymous transversion from } \\ \text{codon \ensuremath{i} to \ensuremath{j} } } \\  
\kappa\cdot\pi_{j} & \substack{ \text{ synonymous transition } } \\
\omega\cdot\pi_{j} & \substack{ \text{ non-synonymous transversion } } \\
\omega\cdot\kappa\cdot\pi_{j} & \substack{ \text{ non-synonymous transversion } } \\
0 & \substack{ \text{ otherwise} }
\end{cases},
\label{eq:gy94}
\end{equation}
% % {\footnotesize
% \begin{equation}
% q_{ij}^{GY94}=\begin{cases}
% \pi_{j} & \text{if \ensuremath{i\rightarrow j} is a synonymous transversion}\\
% \kappa\cdot\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a synonymous transition}\\
% \omega\cdot\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a non-synonymous transversion}\\
% \omega\cdot\kappa\cdot\pi_{j} & \text{if \ensuremath{i\rightarrow j} is a non-synonymous transversion}\\
% 0 & \text{otherwise}
% \end{cases},
% \label{eq:gy94}
% \end{equation}
% % } % END: size

@@

\begin{equation}
\left\{ \mathbf{Q}\right\} _{ij}=\begin{cases}
\alpha\cdot\kappa\cdot\pi_{n}^{p} & \substack{ \text{ \ensuremath{i\rightarrow j} is a synonymous transition from } \\ \text{nucleotide \ensuremath{m} to \ensuremath{n} at codon position \ensuremath{p} } } \\
\alpha\cdot\pi_{n}^{p} & \substack{ \text{ synonymous transversion from } \\ \text{\ensuremath{m} to \ensuremath{n} at position \ensuremath{p} } } \\
\beta\cdot\kappa\cdot\pi_{n}^{p} & \substack{ \text{ non-synonymous transition from } \\ \text{\ensuremath{m} to \ensuremath{n} at position \ensuremath{p} } } \\
\beta\cdot\pi_{n}^{p} & \substack{ \text{ non-synonymous transversion from } \\ \text{\ensuremath{m} to \ensuremath{n} at position \ensuremath{p} } } \\
0 & \substack{ \text{ otherwise}}
\end{cases}.
\label{eq:mg94}
\end{equation}
